Accessing the nucleon transverse structure in inclusive deep inelastic scattering

### 2017-07-25 1607 - JMC for g2/Paper/JMC_dis.tex
Draft mods in v1, after referee report

%%%%%% for draft
\newcommand{\todo}[1]{\marginpar{$\bullet$}\textbf{#1}}
\def\AAcom#1{{\bf  \textcolor{Red}{[AA: {#1}]}}}
\def\AAmod#1{{\textcolor{Green}{#1}}}

@@

\begin{abstract}
We revisit the standard analysis of inclusive Deep Inelastic Scattering
off nucleons taking into  account the fact that on-shell quarks cannot be present in the final state, but they rather decay into hadrons --  a process that can be described in terms of suitable ``jet'' correlators. As a consequence, a spin-flip term
associated with the invariant mass of the produced hadrons is generated
non perturbatively and couples to the target's transversity distribution
function. In inclusive cross sections, this provides an hitherto neglected and
large contribution to the twist-3 part of the $g_2$ structure function, that
can explain the discrepancy between recent calculations and fits of this
quantity.  It also provides an extension of the Burkhardt--Cottingham sum rule, putting constraints on the small-$x$ behavior of the transversity function, as well as an extension of the Efremov--Teryaev--Leader sum rule, suggesting a novel way to measure the tensor charge of the proton.
\end{abstract}

@@

The jet correlators describe interactions of a perturbative quark with vacuum fields, that break chiral symmetry and generate a non perturbative mass of order 100 MeV, much larger than the current quark mass for light flavors, as also heuristically advocated in Ref.~\cite{Afanasev:2007ii} for a study of transverse target single-spin asymmetries in two-photon exchange processes. Here, we formalize this idea in the context of collinear factorization, and observe that jet correlators introduce a new
contribution already in one-photon exchange processes, and more precisely to the inclusive $g_2$ structure function. The new term is proportional to the transversity distribution function multiplied by a new nonperturbative ``jet mass'', which will be precisely defined below, and has the interesting features that:
(a) it violates the Wandzura--Wilczek relation;
(b) it extends the Burkhardt--Cottingham sum rule, providing a useful bound on the small-$x$ behavior of the transversity distribution;
(c) it also extends the Efremov--Teryaev--Leader sum rule, providing a novel way to measure the proton's tensor charge.
We estimate this new jet-mass-induced contribution based on a recent extraction of the transversity distribution, and show it can indeed be very large.

@@

When inserting the jet correlator in the handbag diagram for inclusive DIS, the integration over $dl^+$, or equivalently $dl^2/(2l^-)$, is kinematically coupled to the other integrations, and induces corrections of order ${\cal O}(1/Q^2)$ whose effect on the $F_2$ structure function has been studied in Ref.~\cite{Accardi:2008ne}. In this paper, where we limit our
attention to effects of order ${\cal O}(1/Q)$, we can neglect $k^-$ compared to $q^-$. As a consequence, we can extend the integration over $dl^2$ to infinity, with the consequence that the jet correlator decouples from the parton correlator $\Phi$, and the inclusive structure functions only depend on the integrated jet correlator 
\begin{equation} 
  \Xi(l^-,{\bm l_T}) \equiv \int \frac{dl^2}{2l^-} \, \Xi(l) 
    =  \frac{\Lambda}{2 l^-}\,\xi_1 \id
    +  \xi_2 \frac{\nslash_-}{2} 
    + {\cal O}(l_T/l^-) + \text{higher\ twists} \ .
\end{equation} 
The neglected ${\bm l_T}$-dependent and higher twist terms only contribute to ${\cal O}(1/Q^2)$ to the inclusive cross section. Note that thanks to Eq.~\eqref{eq:jetfnsprops} we obtain
\begin{align}
\xi_1 &= \int d\mu^2 \frac{\mu}{\Lambda} J_1(\mu^2) 
       \equiv \frac{\mj}{\Lambda},
&
\xi_2 &= \int d\mu^2 J_2(\mu^2) = 1 \ ,
\end{align} 
where $\mj$ can be interpreted as the average invariant mass produced in the spin-flip fragmentation processes of a quark of flavor $q$.
It is important to notice that while $\xi_2=1$ exactly due to CPT invariance
(see Sec.~10.7 of Ref.~\cite{Weinberg:1995mt}), the jet mass $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is dynamically determined. From the analytic properties of the spectral functions we know that $J_2(\mu^2) = Z \delta(\mu^2-m_q) + \bar J_2(\mu^2) \theta (\mu^2-m_\pi^2)$ with $Z<1$ and the continuum starting at $m_\pi$ (the mass of the pion) due to color confinement effects. Taking into account that $J_1 < J_2$, we may therefore estimate 
\begin{align}
  \label{eq:mjet}
  \mj = {\cal O}(100 \text{\ MeV}) \ .
\end{align}
Although $\mj$ is in general a nonperturbative quantity, it is interesting to
notice that on the perturbative vacuum, 
\begin{align*}
  \Xi^\text{pert}(l) = (\lslash + \mq \id) \, \delta(l^2-m_q^2) + {\cal O}(\alpha_s) \ ,
\end{align*}
where $\mq$ is the current quark mass; therefore $\mj^\text{pert}=m_q$, and one recovers the result of the calculation with the conventional handbag diagram.
However, we are here considering non perturbative effects
in the quark propagation, and $\mj \gg \mq$. 
Therefore, differently from $J_2$, the $J_1$ function leaves an imprint on the inclusive DIS cross section even in the asymptotic $Q^2 \to \infty$ regime.

@@

The role of the $\xi_1=M_q/\Lambda$ term in inclusive DIS can be discussed by analyzing the following terms of the semi-inclusive hadronic tensor \cite{Mulders:1995dh}:
\begin{align}
  \label{eq:Wsidis_ini}
  2 M  W^{\mu\nu}
    & = i \frac{2M}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} \\
    & \times \sum_q e_q^2
    \bigg[ 2 \xbj g_T^q(\xbj) \sum_h \int d^2p_{hT} dz \, z\, D_1^{q,h}(z,p_{hT}) 
  + \frac{2\Lambda}{M} h_1^q(\xbj) \sum_h \int d^2p_{hT}  dz \, \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots \ , \nonumber
\end{align}
where $g_T^q(z,p_{hT})$ and $\tilde E^q(z,p_{hT})$ are twist-3 TMDs originating, respectively, from the quark-quark and the quark-gluon-quark fragmentation correlators. Note that in Eq.~\eqref{eq:Wsidis_ini} $M$ is the proton's mass, and we identified the power counting scale $M_h$ of Ref.~\cite{Mulders:1995dh} with our $\Lambda$. For clarity, we also reintroduced the quark flavors $q$, $e_q$ being their respective electric charge. 
The first term can be easily integrated with the help of the sum rules
\eqref{eq:SIDIS_to_DIS_TMDlevel1} and \eqref{eq:SIDIS_to_DIS_TMDlevel2}. To integrate the second term, however, we first need make
use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a
consequence of the QCD equations of motion \cite{Bacchetta:2006tn}, then
make again use of the sum rules \eqref{eq:SIDIS_to_DIS_TMDlevel1}-\eqref{eq:SIDIS_to_DIS_TMDlevel2} to obtain
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda }\ .
\end{align}
This formula provides us with a
non perturbative generalization of the commonly used $\int\tilde E =0$ sum
rule introduced in \cite{Jaffe:1996zw}. Indeed, calculating the jet correlator 
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the integral would vanish.

@@

\section{The $g_2$ structure function}

@@

As one can see, in the proton case the pure twist-3 contribution is quite
smaller in magnitude than, and nearly opposite in sign compared to, the twist-3 term extracted in the JAM15 fit. The quark-mass contribution, as expected, is essentially negligible. 
For what concerns the jet-mass contribution, the uncertainties due to the $h_1$
extraction are very large, especially at low $\xbj$. In addition, there is an
overall normalization uncertainty due to the choice of $\mj$, not shown in the
plot. In any case, the jet-mass contribution is strikingly large, and of the same order of magnitude as the chiral-even twist 3 term. If we assume the latter to be of the order of the model calculation
by Braun et al., the breaking of the Wandzura-Wilczek relation can be used to
constrain the extractions of the transversity distribution. This is in particular true at low $\xbj$, where the pure twist-3 term is expected to vanish. 
Moreover, it is quite clear that the gap
between the pure twist-3 $g_2^{\rm tw3}$ function and the JAM15 fit can be
explained by the new jet-mass contribution we discuss in this paper.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\includegraphics[width=8cm]{g2contribN}
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left)
and neutron (right) $g_2$ structure functions compared to the JAM15 fit of the
$g_2-g_2^{\text{WW}}$ function (solid black) \cite{Sato:2016tuz}. The quark and jet
contributions are shown with a dotted red and a dot-dashed green line
respectively, with uncertainty bands coming form the Pavia15 fit of the
transversity function \cite{Radici:2015mwa}. The uncertainty in the choice
$m_q=5$ MeV and $\mj=100$ MeV is not shown. The pure twist-3 contribution
calculated by Braun et al. \cite{Braun:2011aw} is shown as a dashed blue line
(no uncertainty estimate was provided in the original reference). 
}
\end{center}
\end{figure}

@@

One should notice that since $h_1$ is slowly driven to 0 by QCD evolution as $Q^2 \to \infty$, the BC sum rule is still satisfied at least
asymptotically.
At finite scales, however, the only way to preserve the validity of the
BC sum rule is if
$
%\begin{align}
   \int dx\, \frac{1}{x} h_1^q(x) = 0 \ .
\label{eq:ABsumrule}
%\end{align}
$
Interestingly, one can show that this constraint, if valid at any given scale
$Q_0$ is conserved through QCD evolution. However, we think that it is unlikely to be satisfied in general, since it is explicitly broken in perturbative QCD~\cite{Kundu:2001pk}, as well as in model calculations \cite{Schweitzer:2001sr,Wakamatsu:2007nc,Pasquini:2005dk,Cloet:2007em,Bacchetta:2008af,Bourrely:2010ng}. Then, if we assume that the BC sum rule is broken by a {\em finite} amount we obtain that $h_1(x)/x$ must be integrable, which implies the following bound on the small $x$ behavior of the transversity: 
\begin{align}
  h_1^q(x) \propto x^\epsilon \ \ \ \epsilon>0 \ .
\label{eq:ABbound}
\end{align}
This bound can be very useful, \eg, in transversity fits, where the data at
small $x$ is, as yet, very limited, and in general for proper extrapolations
to $x_B=0$ when experimentally measuring moments. We also note that the small-$x$ behavior of the longitudinal spin structure function $g_1$ has been recently investigated in Ref.~\cite{Kovchegov:2016zex}; it would be interesting to extend those techniques to the transverse spin structure functions $g_T=g_1+g_2$ and $h_1$.

@@

For both the $d_1$ and $d_2$ moments, the transversity contribution is a background to the extraction of the pure twist-3 piece. Fortunately, it is a quantity that can be extracted from the lattice
\cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015owa,Bhattacharya:2016zcn} 
or fitted \cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa}.
Furthermore, the bound
\eqref{eq:ABbound} promises to improve future transversity fits. Therefore the pure twist-3 part can, in principle, be properly isolated and measured.

We should also note that the $\mj$ jet mass parameter can
also be experimentally measured, \eg, in electron-positron collisions.
A promising avenue is through inclusive single hadron production, $e^+ e^- \to h X$, and inclusive dihadron 
production from the same hemisphere, $e^+ e^- \to h h X$, see~Fig.\ref{fig:epl_emn_h}.
In single-hadron production, the fragmentation functions 
play the role of PDFs in DIS and couple to the jet functions
in an analogous way. To access the spin-flip $J_1$ function one needs to detect a polarized hadron, such as a $\Delta$ baryon. 
In double hadron production, the enlarged number of Dirac structures
of the dihadron fragmentation correlators related to the relative momentum of the two hadrons \cite{Bacchetta:2002ux,Bacchetta:2003vn} allows one to access
the jet function in novel ways, and in particular to isolate the
contribution from the helicity-flip $J_1$ term in combination with the
chiral-odd fragmentation function $H_1^{\sphericalangle}$.

@@

In this paper, we revisited the inclusive DIS analysis, including the
effects due to the production of a system of final state hadrons in the current direction, which we
conveniently referred to as a ``jet.'' We described this in terms of a jet
correlator that corresponds, up to twist-4 contributions, to the nonperturbative quark cut propagator, or, equivalently, to the quark's spectral function, and of a quark-gluon-quark jet correlator needed to insure gauge invariance of the calculation.
We then carried out the analysis of
the DIS cross section up to contribution of order $1/Q$. 
The introduction of the jet correlators 
leads to a difference in the expression of the structure function $g_2$ in
inclusive DIS with respect to the standard analysis: a new term appears, proportional to a jet mass parameter $\mj={\cal O}$(100 MeV)
and to the transversity distribution function. This new term
contributes to the violation of the Wandzura-Wilczek relation, in addition to
the standard pure twist-3 terms and quark mass corrections. Contrary to these
standard terms, however, the new jet mass correction does not necessarily integrate to zero and so violates also the
Burkhardt--Cottingham and Efremov--Teryaev--Leader sum rules. This is yet another example of how surprising
and rich the phenomenology of polarized inclusive DIS can be.

>>>>

%%%%%% for draft
\newcommand{\todo}[1]{\marginpar{$\bullet$}\textbf{#1}}
\def\AAcom#1{{\bf  \textcolor{Red}{[AA: {#1}]}}}
\def\AAmod#1{{\textcolor{Red}{#1}}}

@@

\begin{abstract}
We revisit the standard analysis of inclusive Deep Inelastic Scattering
off nucleons taking into  account the fact that on-shell quarks cannot be present in the final state, but they rather decay into hadrons --  a process that can be described in terms of suitable ``jet'' correlators. As a consequence, a spin-flip term
associated with the invariant mass of the produced hadrons is generated
nonperturbatively and couples to the target's transversity distribution
function. In inclusive cross sections, this provides an hitherto neglected and
large contribution to the twist-3 part of the $g_2$ structure function, that
can explain the discrepancy between recent calculations and fits of this
quantity.  It also provides an extension of the Burkhardt--Cottingham sum rule, \AAmod{
  %putting constraints
  providing new information}
on the small-$x$ behavior of the transversity function, as well as an extension of the Efremov--Teryaev--Leader sum rule, suggesting a novel way to measure the tensor charge of the proton.
\end{abstract}

@@

The jet correlators describe interactions of a perturbative quark with vacuum fields, that break chiral symmetry and generate a nonperturbative mass
\AAmod{
  % of order 100 MeV
  estimated in the 10-100 MeV range},
much larger than the current quark mass for light flavors, as also heuristically advocated in Ref.~\cite{Afanasev:2007ii} for a study of transverse target single-spin asymmetries in two-photon exchange processes. Here, we formalize this idea in the context of collinear factorization, and observe that jet correlators introduce a new
contribution already in one-photon exchange processes, and more precisely to the inclusive $g_2$ structure function. The new term is proportional to the transversity distribution function multiplied by a new nonperturbative ``jet mass'', which will be precisely defined below, and has the interesting features that:
(a) it violates the Wandzura--Wilczek relation;
(b) it extends the Burkhardt--Cottingham sum rule, providing
\AAmod{
  %a useful bound
  useful information}  
on the small-$x$ behavior of the transversity distribution;
(c) it also extends the Efremov--Teryaev--Leader sum rule, providing a novel way to measure the proton's tensor charge.
We estimate this new jet-mass-induced contribution based on a recent extraction of the transversity distribution, and show it can indeed be very large.

@@

When inserting the jet correlator in the handbag diagram for inclusive DIS, the integration over $dl^+$, or equivalently $dl^2/(2l^-)$, is kinematically coupled to the other integrations, and induces corrections of order ${\cal O}(1/Q^2)$ whose effect on the $F_2$ structure function has been studied in Ref.~\cite{Accardi:2008ne}. In this paper, where we limit our
attention to effects of order ${\cal O}(1/Q)$, we can neglect $k^-$ compared to $q^-$. As a consequence, we can extend the integration over $dl^2$ to infinity, with the consequence that the jet correlator decouples from the parton correlator $\Phi$, and the inclusive structure functions only depend on the integrated jet correlator 
\begin{equation} 
  \Xi(l^-,{\bm l_T}) \equiv \int \frac{dl^2}{2l^-} \, \Xi(l) 
    =  \frac{\Lambda}{2 l^-}\,\xi_1 \id
    +  \xi_2 \frac{\nslash_-}{2} 
    + {\cal O}(l_T/l^-) + \text{higher\ twists} \ .
\end{equation} 
The neglected ${\bm l_T}$-dependent and higher twist terms only contribute to ${\cal O}(1/Q^2)$ to the inclusive cross section. Note that thanks to Eq.~\eqref{eq:jetfnsprops} we obtain
\begin{align}
\xi_1 &= \int d\mu^2 \frac{\mu}{\Lambda} J_1(\mu^2) 
       \equiv \frac{\mj}{\Lambda},
&
\xi_2 &= \int d\mu^2 J_2(\mu^2) = 1 \ ,
\end{align} 
where $\mj$ can be interpreted as the average invariant mass produced in the spin-flip fragmentation processes of a quark of flavor $q$.

\AAmod{
% It is important to notice that while $\xi_2=1$ exactly due to CPT invariance
% (see Sec.~10.7 of Ref.~\cite{Weinberg:1995mt}), the jet mass
% $0 < \mj < \int d\mu^2 \mu J_2(\mu^2)$ is dynamically determined. From the analytic properties of the spectral functions we know that $J_2(\mu^2) = Z \delta(\mu^2-m_q) + \bar J_2(\mu^2) \theta (\mu^2-m_\pi^2)$ with $Z<1$ and the continuum starting at $m_\pi$ (the mass of the pion) due to color confinement effects. Taking into account that $J_1 < J_2$, we may therefore estimate 
% \begin{align}
%   \label{eq:mjet}
%   \mj = {\cal O}(100 \text{\ MeV}) \ .
% \end{align}
It is important to notice that, while $\xi_2=1$ exactly due to CPT invariance
(see Sec.~10.7 of Ref.~\cite{Weinberg:1995mt}), the jet mass
$\mj < \int d\mu^2 \mu J_2(\mu^2)$ is dynamically determined. From the analytic properties of the spectral functions we know that $J_2(\mu^2) = Z \delta(\mu^2-m_q) + \bar J_2(\mu^2) \theta (\mu^2-m_\pi^2)$ with $Z<1$ and the continuum starting at $m_\pi$ (the mass of the pion) due to color confinement effects, indicating $M_q=O(\Lambda_{QCD})$. However, in a dynamical confinement scenario, the spectral function $J_2$ needs not be positive definite \cite{Roberts:2016vyn} and we therefore estimate
\begin{align}
  M_q\sim 10-100 \text{\ MeV}.
\end{align}
An experimental measurement of $M_q$ is anyway possible, as we discuss in Sections~\ref{sec:moments}, and could shed light on the confinement mechanism. We have also explicitly verified that $M_q>m_q$ in a model where quark fragmentation is simulated by a Yukawa pseudoscalar quark-meson interaction.
}

Although $\mj$ is in general a nonperturbative quantity, it is interesting to
notice that on the perturbative vacuum, 
\begin{align*}
  \Xi^\text{pert}(l) = (\lslash + \mq \id) \, \delta(l^2-m_q^2) + {\cal O}(\alpha_s) \ ,
\end{align*}
where $\mq$ is the current quark mass; therefore $\mj^\text{pert}=m_q$, and one recovers the result of the calculation with the conventional handbag diagram.
However, we are here considering nonperturbative effects
in the quark propagation, and $\mj \gg \mq$. 
Therefore, differently from $J_2$, the $J_1$ function leaves an imprint on the inclusive DIS cross section even in the asymptotic $Q^2 \to \infty$ regime.

@@

The role of the $\xi_1=M_q/\Lambda$ term in inclusive DIS can be discussed by analyzing the following terms of the semi-inclusive hadronic tensor \cite{Mulders:1995dh}:
\begin{align}
  \label{eq:Wsidis_ini}
  2 M  W^{\mu\nu}
    & = i \frac{2M}{Q} \hat t^{[\mu}_{\phantom \perp} 
    \epsilon_\perp^{\nu]\rho}S_{\perp\rho} \\
    & \times \sum_q e_q^2
    \bigg[ 2 \xbj g_T^q(\xbj) \sum_h \int d^2p_{hT} dz \, z\, D_1^{q,h}(z,p_{hT}) 
  + \frac{2\Lambda}{M} h_1^q(\xbj) \sum_h \int d^2p_{hT}  dz \, \tilde E^{q,h}(z,p_{hT}) \bigg] + \ldots \ , \nonumber
\end{align}
where $g_T^q(z,p_{hT})$ and $\tilde E^q(z,p_{hT})$ are twist-3 TMDs originating, respectively, from the quark-quark and the quark-gluon-quark fragmentation correlators. Note that in Eq.~\eqref{eq:Wsidis_ini} $M$ is the proton's mass, and we identified the power counting scale $M_h$ of Ref.~\cite{Mulders:1995dh} with our $\Lambda$. For clarity, we also reintroduced the quark flavors $q$, $e_q$ being their respective electric charge. 
The first term can be easily integrated with the help of the sum rules
\eqref{eq:SIDIS_to_DIS_TMDlevel1} and \eqref{eq:SIDIS_to_DIS_TMDlevel2}. To integrate the second term, however, we first need make
use of the relation $\tilde E(z) = E(z) - (\mq/\Lambda) z D_1(z)$, which is a
consequence of the QCD equations of motion \cite{Bacchetta:2006tn}, then
make again use of the sum rules \eqref{eq:SIDIS_to_DIS_TMDlevel1}-\eqref{eq:SIDIS_to_DIS_TMDlevel2} to obtain
\begin{align}
  \sum_h \int dz d^2p_{hT} \tilde E^{q,h}(z,p_{hT}) 
    = \xi_1 - \frac{\mq}{\Lambda} \xi_2 = \frac{\mj - \mq}{\Lambda }\ .
\end{align}
This formula provides us with a
nonperturbative generalization of the commonly used $\int\tilde E =0$ sum
rule introduced in \cite{Jaffe:1996zw}. Indeed, calculating the jet correlator 
on the perturbative vacuum one would obtain, as already discussed, $\mj=\mq$
and the integral would vanish.

@@

\section{The $g_2$ structure function}
\label{sec:g2}

@@

As one can see, in the proton case the pure twist-3 contribution is quite
smaller in magnitude than, and nearly opposite in sign compared to, the twist-3 term extracted in the JAM15 fit. The quark-mass contribution, as expected, is essentially negligible. 
For what concerns the jet-mass contribution, the uncertainties due to the $h_1$
extraction are very large, especially at low $\xbj$. In addition, there is an
overall normalization uncertainty due to the choice of $\mj$, not shown in the
plot. In any case, the jet-mass contribution is strikingly large, and of the same order of magnitude as the chiral-even twist 3 term.

If we assume the latter to be of the order of the model calculation
by Braun et al., the breaking of the Wandzura-Wilczek relation can be used to
constrain the extractions of the transversity distribution. This is in particular true at low $\xbj$, where the pure twist-3 term is expected to vanish. 
Moreover, it is quite clear that the gap
between the pure twist-3 $g_2^{\rm tw3}$ function and the JAM15 fit can be
explained by the new jet-mass contribution we discuss in this paper.

@@

\begin{figure}[tb]
\begin{center}
\includegraphics[width=8cm]{g2contrib}
\includegraphics[width=8cm]{g2contribN}
\vskip-0.2cm
\caption{\label{f:g2contrib} 
Different contributions to the non Wandzura-Wilczek part of the proton (left)
and neutron (right) $g_2$ structure functions compared to the JAM15 fit of the
$g_2-g_2^{\text{WW}}$ function (solid black) \cite{Sato:2016tuz}. The quark and jet
contributions are shown with a dotted red and a dot-dashed green line
respectively, with uncertainty bands coming form the Pavia15 fit of the
transversity function \cite{Radici:2015mwa}. The uncertainty in the choice
$m_q=5$ MeV and $\mj=100$ MeV is not shown. The pure twist-3 contribution
calculated by Braun et al. \cite{Braun:2011aw} is shown as a dashed blue line
(no uncertainty estimate was provided in the original reference). 
}
\end{center}
\end{figure}

@@

One should notice that since $h_1$ is slowly driven to 0 by QCD evolution as $Q^2 \to \infty$, the BC sum rule
\AAmod{
  %is still
  may still be }
satisfied at least asymptotically.
At finite scales, however, the only way to preserve the validity of the
BC sum rule is if
$
%\begin{align}
   \int dx\, \frac{1}{x} h_1^q(x) = 0 \ .
\label{eq:ABsumrule}
%\end{align}
$
Interestingly, one can show that this constraint, if valid at any given scale
$Q_0$ is conserved through QCD evolution. However, we think that it is unlikely to be satisfied in general, since
\AAmod{
  %it is explicitly broken
  the right hand side is different from zero}
in perturbative QCD~\cite{Kundu:2001pk}, as well as in model calculations \cite{Schweitzer:2001sr,Wakamatsu:2007nc,Pasquini:2005dk,Cloet:2007em,Bacchetta:2008af,Bourrely:2010ng}.
\AAmod{
% Then, if we assume that the BC sum rule is broken by a {\em finite} amount we obtain that $h_1(x)/x$ must be integrable, which implies the following bound on the small $x$ behavior of the transversity: 
% \begin{align}
%   h_1^q(x) \propto x^\epsilon \ \ \ \epsilon>0 \ .
% \label{eq:ABbound}
% \end{align}
% This bound can be very useful, \eg, in transversity fits, where the data at
% small $x$ is, as yet, very limited, and in general for proper extrapolations
% to $x_B=0$ when experimentally measuring moments.
A finite breaking of the BC sum rule would imply that $h_1(x)/x$ must be integrable, which is possible only if, at small $x$, the transversity goes as $h_1^q(x) \propto x^\epsilon$ with $\epsilon>0$. While $\epsilon=1$ in perturbative QCD \cite{Hayashigaki:1997dn,Vogelsang:1997ak}, the leading Regge contribution at small $x$ indicates that $\epsilon = 0$  \cite{Kirschner:1996jj}, and opens the door to a more serious breaking of the sum rule.}
Finally, we note that the small-$x$ behavior of the longitudinal spin structure function $g_1$ has been recently investigated in Ref.~\cite{Kovchegov:2016zex}; \AAmod{
however, since the small $x$ structure of the $h_1$ function may be quite different from that of $g_1$ \cite{Kirschner:1996jj},}
it would be interesting to extend those techniques to the transverse spin structure functions $g_T=g_1+g_2$ and $h_1$.

@@

For both the $d_1$ and $d_2$ moments, the transversity contribution is a background to the extraction of the pure twist-3 piece. Fortunately, it is a quantity that can be extracted from the lattice
\cite{Green:2012ej,Bali:2014nma,Bhattacharya:2015wna,Abdel-Rehim:2015owa,Bhattacharya:2016zcn} 
or fitted \cite{Radici:2015mwa,Anselmino:2015sxa,Kang:2015msa},
\AAmod{
% Furthermore, the bound
% \eqref{eq:ABbound} promises to improve future transversity fits. Therefore the pure twist-3 part can, in principle, be properly isolated and measured.
and information from the extended BC sum rule \eqref{eq:BC-sfn} promises to improve future transversity fits. Furthermore, as combined QCD fits of different distribution function have now become possible \cite{Ethier:2017zbq}, the jet mass $M_q$ could also be considered as a free parameter in a combined helicity and transversity PDF fit. Therefore the pure twist-3 part can, in principle, be properly isolated and measured.
}

We should also note that the $\mj$ jet mass parameter can
be experimentally measured, \eg, in electron-positron collisions.
A promising avenue is through inclusive single hadron production, $e^+ e^- \to h X$, and inclusive dihadron 
production from the same hemisphere, $e^+ e^- \to h h X$, see~Fig.~\ref{fig:epl_emn_h}.
In single-hadron production, the fragmentation functions 
play the role of PDFs in DIS and couple to the jet functions
in an analogous way. To access the spin-flip $J_1$ function one needs to detect a polarized hadron, such as a $\Delta$ baryon. 
In double hadron production, the enlarged number of Dirac structures
of the dihadron fragmentation correlators related to the relative momentum of the two hadrons \cite{Bacchetta:2002ux,Bacchetta:2003vn} allows one to access
the jet function in novel ways, and in particular to isolate the
contribution from the helicity-flip $J_1$ term in combination with the
chiral-odd fragmentation function $H_1^{\sphericalangle}$.

@@

In this paper, we revisited the inclusive DIS analysis, including the
effects due to the production of a system of final state hadrons in the current direction, which we
conveniently referred to as a ``jet.'' We described this in terms of a jet
correlator that corresponds, up to twist-4 contributions, to the nonperturbative quark cut propagator, or, equivalently, to the quark's spectral function, and of a quark-gluon-quark jet correlator needed to insure gauge invariance of the calculation.
We then carried out the analysis of
the DIS cross section up to contribution of order $1/Q$. 
The introduction of the jet correlators 
leads to a difference in the expression of the structure function $g_2$ in
inclusive DIS with respect to the standard analysis: a new term appears, proportional to a jet mass parameter
\AAmod{ $\mj={\cal O}$(10-100 MeV) }
and to the transversity distribution function. This new term
contributes to the violation of the Wandzura-Wilczek relation, in addition to
the standard pure twist-3 terms and quark mass corrections. Contrary to these
standard terms, however, the new jet mass correction does not necessarily integrate to zero and so violates also the
Burkhardt--Cottingham and Efremov--Teryaev--Leader sum rules. This is yet another example of how surprising
and rich the phenomenology of polarized inclusive DIS can be,
\AAmod{that offers a new direction for theoretical studies and experimental investigations of spin physics over a wide range in $x$, from the valence and sea regions at Jefferson Lab \cite{Dudek:2012vr} to the small-$x$ region at the future Electron-Ion Collider \cite{Accardi:2012qut}.}
